Modularly Composing Typed Language Fragments (Supplemental Material)

### 2014-11-16 att-pldi15/att-pldi15-supplement.tex
final PLDI15 submission

%\documentclass[12pt]{article}
\documentclass[10pt,preprint]{sigplanconf}
% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.
\usepackage{amsmath}
\usepackage{amssymb} 
\usepackage{amsthm}
\usepackage{ stmaryrd }
\usepackage{mathpartir}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\input{macros-catlam}

@@

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{tyconinvariant}{Tycon Invariant}
\newenvironment{proof-sketch}{\noindent{\emph{Proof Sketch.}}}{\qed}
\makeatletter

@@

@@

\section{Internal Language}
We assume the statics of the IL are specified in the standard way by judgements for  type formation {$\iDelta \vdash \tau$}, typing context formation { $\iDelta \vdash \iGamma$} and type assignment {$\iDelta~\iGamma \vdash \iota : \tau\moutput$}. 
%The typing and type formation contexts obey standard structural properties (i.e. weakening, exchange and contraction). 
The internal dynamics are specified as a structural operational semantics with a stepping judgement {\small $\iota \mapsto \iota\moutput$} and a value judgement {$\iota~\mathtt{val}$}. The multi-step judgement $\iota \mapsto^{*} \iota\moutput$ is the reflexive, transitive closure of the stepping judgement and the evaluation judgement $\iota \Downarrow \iota'$ is defined iff $\iota \mapsto^{*} \iota'$ and $\iota'~\mathtt{val}$. Both the static and dynamic semantics of the IL can be found in any standard textbook covering typed lambda calculi (we directly follow \cite{pfpl}), so we assume familiarity and give the lemmas in this section without proof.

We use {$\mathcal{L}\{{\rightharpoonup}\,{\forall}\,{\mu}\,{\iunit}\,{\times}\,{+}\}$}, the syntax for which is shown in Figure \ref{syntax-IL}, as representative of any intermediate language for a typed functional language. In fact, our intention is not to prescribe a particular choice of IL, so we will here only review the key metatheoretic properties that the IL must possess. Each choice of IL is technically a distinct dialect of @$\lambda$, but for the broad class of ILs that enjoy these properties, the metatheory in the remainder of the supplement should follow without trouble.

\begin{lemma}[Internal Type Assignment]
If $\Delta \vdash \Gamma$ and $\Delta~\Gamma \vdash \iota : \tau$ then $\Delta \vdash \tau$.
\end{lemma}

Internal type safety follows the standard methodology of proving preservation and progress lemmas.

\begin{lemma}[Internal Progress]
If $\emptyset~\emptyset \vdash \iota : \tau$ then either $\iota~\texttt{val}$ or $\iota \mapsto \iota'$.
\end{theorem}
\begin{lemma}[Internal Preservation]
If $\emptyset~\emptyset \vdash \iota : \tau$ and $\iota \mapsto \iota'$ then $\emptyset~\emptyset \vdash \iota' : \tau$. 
\end{lemma}

We assume substitution and simultaneous $n$-ary substitutions, $\delta$ and $\gamma$, have the standard semantics, and that typing and type formation contexts, $\Delta$ and $\Gamma$, obey standard properties of weakening, exchange and contraction. We assume that the names of variables and type variables are unimportant, so that $\alpha$-equivalent terms, types and contexts can be identified implicitly throughout this work. In particular, we need the definitions of substitution validity shown in Figures \ref{fig:internal-type-substitution-validity} and \ref{fig:internal-term-substitution-validity} and the following lemmas.

\begin{figure}[t]
\small
\hspace{-5px}\input{syntax-table-IL}
\caption{Syntax of {$\mathcal{L}\{{\rightharpoonup}\,{\forall}\,{\mu}\,{\iunit}\,{\times}\,{+}\}$}, our internal language (IL). Metavariable $x$ ranges over term variables and $\alpha$ and $t$ both range over type variables.}
\label{syntax-IL}
\end{figure}

\begin{figure}
\small\fbox{$\Delta \vdash \delta : \Delta$}~~$\delta ::=  \emptyset ~|~ \delta, \tau/\alpha$

\begin{mathpar}
\inferrule[tysub-emp]{ }{\Delta \vdash \emptyset : \emptyset}

\inferrule[tysub-ext]{
    \Delta \vdash \delta : \Delta'\\
    \Delta \vdash \tau
}{
    \Delta \vdash \delta, \tau/\alpha : \Delta', \alpha
}
\end{mathpar}
\caption{Internal Type Substitution Validity}
\label{fig:internal-type-substitution-validity}
\end{figure}
\begin{figure}
\small\fbox{$\Delta~\Gamma \vdash \gamma : \Gamma$}~~$\gamma ::= \emptyset ~|~ \gamma, \iota/x$
\begin{mathpar}
\inferrule[tmsub-emp]{ }{\Delta~\Gamma \vdash \emptyset : \emptyset}

\inferrule[tmsub-ext]{
    \Delta~\Gamma \vdash \gamma : \Gamma'\\
    \Delta \vdash \tau\\
    \Delta~\Gamma \vdash \iota : \tau
}{
    \Delta~\Gamma \vdash \gamma, \iota/x : \Gamma', x : \tau
}
\end{mathpar}
\caption{Internal Term Substitution Validity}
\label{fig:internal-term-substitution-validity}
\end{figure}
\begin{lemma}[Internal Type Substitution on Types] If $\Delta \vdash \delta : \Delta'$ and $\Delta\Delta' \vdash \tau$ then $\Delta \vdash [\delta]\tau$.\end{lemma}

\begin{lemma}[Internal Type Substitutions on Typing Contexts] If $\Delta \vdash \delta : \Delta'$ and $\Delta\Delta' \vdash \Gamma$ then $\Delta \vdash [\delta]\Gamma$.
\end{lemma}

\begin{lemma}[Internal Type Substitutions on Terms] If $\Delta \vdash \delta : \Delta'$ and $\Delta\Delta' \vdash \tau$ and $\Delta\Delta' \vdash \Gamma$ and $\Delta\Delta'~\Gamma \vdash \iota : \tau$ then $\Delta~[\delta]\Gamma \vdash [\delta]\iota : [\delta]\tau$.
\end{lemma}

\begin{lemma}[Internal Term Substitutions] If $\Delta \vdash \Gamma$ and $\Delta \vdash \Gamma'$ and $\Delta~\Gamma \vdash \gamma : \Gamma'$ and $\Delta~\Gamma\Gamma' \vdash \iota : \tau$ then $\Delta~\Gamma \vdash [\gamma]\iota : \tau$.
\end{lemma}

\section{Tycon Contexts}
\begin{figure}\fbox{$\vdash \Phi$}\vspace{-25px}
\begin{mathpar}
\small
\inferrule[tcc-emp]{ }{\vdash \cdot}

\inferrule[tcc-ext]{
    \vdash \Phi\\
    \tc \notin \text{dom}(\Phi)\\
    \keq{\emptyset}{\ktyidx}\\
    \sofkz{\emptyset}{\emptyset}{\Phi}{\stx{schema}}{\ktyidx \rightarrow \kity}\\\\
    \vdash_{\Phi,  \tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}}} \omega \sim \tcsig{\ktyidx}{\chi}
}{
    \vdash \Phi, \tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}}
}
\end{mathpar}
\caption{Tycon context well-definedness.}
\label{fig:tycon-ctxs}
\end{figure}
\begin{figure}[t]
\small\fbox{$\vdash_\Phi \omega \sim \psi$}
\begin{mathpar}\small
\inferrule[ocstruct-intro]{
    \introsig{\klitidx} \in \chi\\
    \emptyset \vdash \klitidx\\\\
    \sofkz{\emptyset}{\emptyset}{\Phi}{\stx{def}}{\ktyidx \rightarrow \klitidx \rightarrow \kargs \rightarrow \kitm}
}{
    \vdash_\Phi \tcstructn{\stx{def}} \sim \tcsig{\ktyidx}{\chi}
}

\inferrule[ocstruct-targ]{
    \vdash_\Phi \omega \sim \tcsig{\ktyidx}{\chi}\\
    \opname{op}\notin\text{dom}(\chi)\\
    \emptyset \vdash \klitidx\\\\
    \sofkn{\emptyset}{\emptyset}{\Phi}{0}{\stx{def}}{\ktyidx \rightarrow \klitidx \rightarrow \kargs \rightarrow (\kty \times \kitm)}
}{
    \vdash_\Phi \tcstructc{\omega}{op}{\stx{def}} \sim \tcsig{\ktyidx}{\chi, \opsig{\opname{op}}{\klitidx}}
}
\end{mathpar}
\caption{Checking opcon structures against tycon signatures.}
\label{ocstruct}
\end{figure}

\subsection{Tycon Context Well-Definedness}

\subsection{Equality Kinds}
Need an equational theory for SL to state equality kind property, but not important for other metatheory.

\subsection{Full Examples}

\section{Static Language}
\subsection{Kind Formation}

\subsection{Kinding Context Formation}

\subsection{Kinding}

\subsection{Dynamic Semantics}

\subsection{Kind Safety}

\section{Types}
\subsection{Type Translations}

\subsection{Typing Context Translations}

\paragraph{Unicity}
The rules are structured so that if a term is well-typed, both its type and translation are unique.\todo{could move this whole thing to supplement if room needed}
\begin{theorem}[Unicity]
If $\vdash \Phi$ and $\vdash_\Phi \Upsilon \leadsto \Gamma$ and $\vdash_\Phi \st \leadsto \tau$ and $\vdash_\Phi \st' \leadsto \tau'$ and $\eanaX{e}{\st}{\iota}$ and $\eanaX{e}{\st'}{\iota'}$ then $\st = \st'$ and $\tau=\tau'$ and $\iota = \iota'$.
\end{theorem}

\section{External Language}
\subsection{Additional Desugarings}
\subsection{Typing}
\subsection{Proof of Regular String Soundness Tycon Invariant}

\newpage
% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrv}

>>>>

%\documentclass[12pt]{article}
\documentclass[12pt]{article}
\usepackage{fullpage,cmu-titlepage2}
\usepackage{times}
% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.
\usepackage{amsmath}
\usepackage{amssymb} 
\usepackage{amsthm}
\usepackage{ stmaryrd }
\usepackage{mathpartir}
\usepackage{enumerate}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{tabularx}
\input{macros-catlam}
\usepackage{morefloats}

@@

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{tyconinvariant}{Tycon Invariant}
\newtheorem{assumption}{Assumption}
\newenvironment{proof-sketch}{\noindent{\emph{Proof Sketch.}}}{\qed}
\makeatletter

@@

\newcommand{\pfcase}[1]{\textbf{Case}~#1. }
\newcommand{\pfcases}[1]{\textbf{Cases}~#1. }

\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}

@@

\tableofcontents 
\listoffigures

\newpage

\section{Internal Language}
\subsection{Semantics}
We assume an internal language where the statics are specified in the standard way by judgements for  type formation {$\iDelta \vdash \tau$}, typing context formation { $\iDelta \vdash \iGamma$} and type assignment {$\iDelta~\iGamma \vdash \iota : \tau\moutput$}. 
%The typing and type formation contexts obey standard structural properties (i.e. weakening, exchange and contraction). 
The internal dynamics are specified as a structural operational semantics with a stepping judgement {\small $\iota \mapsto \iota\moutput$} and a value judgement {$\iota~\mathtt{val}$}. The multi-step judgement $\iota \mapsto^{*} \iota\moutput$ is the reflexive, transitive closure of the stepping judgement and the evaluation judgement $\iota \Downarrow \iota'$ is defined iff $\iota \mapsto^{*} \iota'$ and $\iota'~\mathtt{val}$. Both the static and dynamic semantics of the IL can be found in any standard textbook covering typed lambda calculi (we directly follow \cite{pfpl}), so we assume familiarity and give the lemmas in this section without proof.

Our particular choice in the paper is {$\mathcal{L}\{{\rightharpoonup}\,{\forall}\,{\mu}\,{\iunit}\,{\times}\,{+}\}$}, the syntax for which is shown in Figure \ref{syntax-IL}, as representative of any intermediate language for a typed functional language. In fact, our intention is not to prescribe a particular choice of IL, so we will here only review the key metatheoretic properties that the IL must possess. Each choice of IL is technically a distinct dialect of @$\lambda$, but for the broad class of ILs that enjoy these properties, the metatheory in the remainder of the supplement should follow without trouble.

\begin{assumption}[Internal Type Assignment]
If $\Delta \vdash \Gamma$ and $\Delta~\Gamma \vdash \iota : \tau$ then $\Delta \vdash \tau$.
\end{assumption}

Internal type safety follows the standard methodology: preservation and progress lemmas.

\begin{assumption}[Internal Progress]
If $\emptyset~\emptyset \vdash \iota : \tau$ then either $\iota~\texttt{val}$ or $\iota \mapsto \iota'$.
\end{assumption}
\begin{assumption}[Internal Preservation]
If $\emptyset~\emptyset \vdash \iota : \tau$ and $\iota \mapsto \iota'$ then $\emptyset~\emptyset \vdash \iota' : \tau$. 
\end{assumption}

We assume that typing and type formation contexts, $\Delta$ and $\Gamma$, are identified up to exchange and contraction and the internal semantics obeys weakening. We assume that the names of variables and type variables are unimportant, so that $\alpha$-equivalent terms, types and contexts are identified implicitly throughout this work.

\subsection{Explicit Substitutions}
We also assume substitution is defined in the usual way. We define explicit $n$-ary substitutions for type variables, $\delta$, and term variables, $\gamma$, also in the standard way, writing $[\delta]\Gamma$, $[\delta]\tau$ and $[\delta]\iota$ for application of type substitutions and $[\gamma]\iota$ for application of term substitutions. We need the definitions of substitution validity shown in Figures \ref{fig:internal-type-substitution-validity} and \ref{fig:internal-term-substitution-validity} and assume the following lemmas.

\begin{assumption}[Internal Term Substitutions] 
\label{assumption:internal-term-substituions}
If $\Delta \vdash \Gamma$ and $\Delta \vdash \Gamma'$ and $\Delta~\Gamma \vdash \gamma : \Gamma'$ and $\Delta~\Gamma\Gamma' \vdash \iota : \tau$ then $\Delta~\Gamma \vdash [\gamma]\iota : \tau$.
\end{assumption}

\begin{assumption}[Internal Type Substitution on Types]
\label{lemma:internal-type-substitution-on-types} If $\Delta \vdash \delta : \Delta'$ and $\Delta\Delta' \vdash \tau$ then $\Delta \vdash [\delta]\tau$.\end{assumption}

\begin{assumption}[Internal Type Substitutions on Typing Contexts] 
\label{lemma:internal-type-substitutions-on-typing-contexts}
If $\Delta \vdash \delta : \Delta'$ and $\Delta\Delta' \vdash \Gamma$ then $\Delta \vdash [\delta]\Gamma$.
\end{assumption}

\begin{assumption}[Internal Type Substitutions on Terms] 
\label{lemma:internal-type-substitutions-on-terms}
If $\Delta \vdash \delta : \Delta'$ and $\Delta\Delta' \vdash \tau$ and $\Delta\Delta' \vdash \Gamma$ and $\Delta\Delta'~\Gamma \vdash \iota : \tau$ then $\Delta~[\delta]\Gamma \vdash [\delta]\iota : [\delta]\tau$.
\end{assumption}

\newpage
\section{Tycon Contexts}
Tycon contexts, $\Phi$, are semantically defined as ordered mappings from tycon names $\tc$ to tycon definitions. As with all ordered mappings, we implicitly assume that all names are unique and that certain operations, e.g. $\text{dom}(\Phi)$, which extracts a set of tycon names, and $\tcdef{\tc}{\psi}{\theta} \in \Phi$, which checks whether a tycon definition is in $\Phi$, and $\Phi\Phi'$, which concatenates two tycon contexts with disjoint domains, are available without defining them explicitly.

\subsection{Validity}

The tycon context validity judgement $\vdash \Phi$ is specified in Figure \ref{fig:tycon-ctxs}. The kinding and kind formation rules will be given in the next section.

\begin{lemma}[Tycon Context Validity]
\label{lemma:tycon-context-validity}
If (a) $\vdash \Phi$ and (b) $\tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}} \in \Phi$ then there exists a prefix $\Phi'$ of $\Phi$ such that (i) $\vdash \Phi'$ and (ii) $\vdash \Phi', \tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}}$ and (iii) $\keq{}{\ktyidx}$ and (iv) $\sofkz{\emptyset}{\emptyset}{\Phi'}{\stx{schema}}{\ktyidx \rightarrow \kity}$ and (v) $\vdash_{\Phi',  \tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}}} \omega \sim \tcsig{\ktyidx}{\chi}$.
\end{lemma}
\begin{proof} Rule induction on (a) and (b).

\pfcase{(tcc-emp)} Does not apply by (b).

\pfcase{(tcc-ext)} We have two cases by (b):\begin{quote}
\pfcase{$\Phi=\Phi', \tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}}$} In this case, we pick $\Phi'$ and have our conclusions:
\begin{enumerate}
\item[(i)] $\vdash \Phi'$ (premise 1 of (tcc-ext))
\item[(ii)] $\vdash \Phi', \tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}}$ (by assumption (a))
\item[(iii)] $\keq{}{\ktyidx}$ (premise 2 of (tcc-ext))
\item[(iv)] $\sofkz{\emptyset}{\emptyset}{\Phi'}{\stx{schema}}{\ktyidx \rightarrow \kity}$ (premise 3 of (tcc-ext))
\item[(v)] $\vdash_{\Phi', \tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}}} \omega \sim \tcsig{\ktyidx}{\chi}$ (premise 4 of (tcc-ext))
\end{enumerate}

\pfcase{$\Phi=\Phi', \tcdef{\tc'}{\psi'}{\theta'}$ for $\tc\neq\tc'$ and (1) $\tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}} \in \Phi'$} We have the following:

\begin{enumerate}
\item[(2)] $\vdash \Phi'$ (premise 1 of (tcc-ext))
\item[] Apply IH to (2) and (1). There exists a prefix $\Phi''$ of $\Phi'$ such that 
\item[(3)] $\vdash \Phi''$
\item[(4)] $\vdash \Phi'', \tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}}$
\item[(5)] $\keq{}{\ktyidx}$ 
\item[(6)] $\sofkz{\emptyset}{\emptyset}{\Phi''}{\stx{schema}}{\ktyidx \rightarrow \kity}$
\item[(7)] $\vdash_{\Phi'',  \tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}}} \omega \sim \tcsig{\ktyidx}{\chi}$
\end{enumerate}

By transitivity of prefixing, we have that $\Phi''$ is a prefix of $\Phi$. Picking $\Phi''$, conclusions (i-v) are (3-7), respectively.
\end{quote}
\end{proof}

In the final premise of (tcc-ext), the opcon structure is checked against the tycon's signature by the rules in Figure \ref{ocstruct}. Opcon structures and opcon signatures are finite maps, so we implicitly assume that all operator names are unique and standard operations are defined as above.

\begin{lemma}[Intro Opcon Existence and Validity]
\label{lemma:intro-opcon-existence-and-validity}
If $\vdash_\Phi \omega \sim \tcsig{\ktyidx}{\chi}$ then (i) $\keyw{intro}[\klitidx] \in \chi$ and (ii) $\emptyset \vdash \klitidx$ and (iii) $\keyw{ana~intro}=\stx{def}\in\omega$ and (iv) $\sofkz{\emptyset}{\emptyset}{\Phi}{\stx{def}}{\ktyidx \rightarrow \klitidx \rightarrow \kargs \rightarrow \kitm}$.
\end{lemma}
\begin{proof}
Rule induction on assumption.

\pfcase{(ocstruct-intro)} We have that $\omega=\keyw{ana~intro}=\stx{def}$ and $\chi=\keyw{intro}[\klitidx]$.

\begin{enumerate}[(i)]
\item $\keyw{intro}[\klitidx] \in \keyw{intro}[\klitidx]$ (by definition)
\item $\emptyset \vdash \klitidx$  (premise 1)
\item  $\keyw{ana~intro}=\stx{def}\in\keyw{ana~intro}=\stx{def}$ (by definition)
\item $\sofkz{\emptyset}{\emptyset}{\Phi}{\stx{def}}{\ktyidx \rightarrow \klitidx \rightarrow \kargs \rightarrow \kitm}$ (premise 2)
\end{enumerate}

\pfcase{(ocstruct-targ)} We have that $\omega=\tcstructc{\omega'}{op}{\stx{def}}$ and  $\chi=\chi'; \opsig{\opname{op}}{\klitidx'}$.
\begin{enumerate}[(1)]
\item $\vdash_\Phi \omega' \sim \tcsig{\ktyidx}{\chi}$ (premise 1)
\item[] Apply IH to (1):
\item $\introsig{\klitidx} \in \chi'$
\item $\emptyset \vdash \klitidx$
\item $\keyw{ana~intro}=\stx{def}\in\omega'$
\item $\sofkz{\emptyset}{\emptyset}{\Phi}{\stx{def}}{\ktyidx \rightarrow \klitidx \rightarrow \kargs \rightarrow \kitm}$
\end{enumerate}
We can now conclude:
\begin{enumerate}[(i)]
\item $\introsig{\klitidx} \in \chi$ by extension of finite mappings on (2)
\item $\emptyset \vdash \klitidx$ by (3)
\item $\keyw{ana~intro}=\stx{def}\in\omega$ by extension of finite mappings on (4)
\item $\sofkz{\emptyset}{\emptyset}{\Phi}{\stx{def}}{\ktyidx \rightarrow \klitidx \rightarrow \kargs \rightarrow \kitm}$ by (5).
\end{enumerate}
\end{proof}
\begin{lemma}[Targeted Opcon Validity]
\label{lemma:targeted-opcon-validity}
If (a) $\vdash_\Phi \omega \sim \tcsig{\ktyidx}{\chi}$ and (b) $\keyw{syn}~\opname{op}=\stx{def} \in \omega$ then (i) $\opname{op}[\klitidx] \in \chi$ and (ii) $\emptyset \vdash \klitidx$ and (iii) $
    \sofkn{\emptyset}{\emptyset}{\Phi}{0}{\stx{def}}{\ktyidx \rightarrow \klitidx \rightarrow \kargs \rightarrow (\kty \times \kitm)}$.
\end{lemma}
\begin{proof}
Rule induction on assumption (a).

\pfcase{(ocstruct-intro)} Does not apply by (b).

\pfcase{(ocstruct-targ)} We have two cases by (b).\begin{quote}\pfcase{$\omega=\omega'; \keyw{syn}~\opname{op}=\stx{def}$ and $\chi=\chi', \opname{op}[{\klitidx}]$} Conclusion (i) follows by definition and conclusions (ii) and (iii) are premises 2 and 3 of (ocstruct-targ).

\pfcase{$\omega=\omega'; \keyw{syn}~\opname{op}'=\stx{def}'$ such that $\opname{op}\neq\opname{op'}$ and (1) $\keyw{syn}~\opname{op}=\stx{def} \in \omega'$} In this case, we have 
    \begin{enumerate}
    \item[(2)] $\vdash_\Phi \omega' \sim \tcsig{\ktyidx}{\chi'}$ (premise 1 of (ocstruct-targ))
    \end{enumerate}

    Applying the IH to (2) and (1), we have:
    \begin{enumerate}[(1)]
    \item[(3)] $\opname{op}[\klitidx] \in \chi'$
    \item[(4)] $\emptyset \vdash \klitidx$
    \item[(5)] $\sofkn{\emptyset}{\emptyset}{\Phi}{0}{\stx{def}}{\ktyidx \rightarrow \klitidx \rightarrow \kargs \rightarrow (\kty \times \kitm)}$
    \end{enumerate}

    We thus have conclusion (i) by extension of finite mappings on (3), and conclusions (ii) and (iii) are (4) and (5).
\end{quote}
\end{proof}

\subsection{Signature Validity}

For the purposes of the SL, only the type signatures are relevant.\footnote{In fact, only the type index kind is relevant, though in future work the remainder will likely be and it is cleaner to define complete signature checking.} It is thus useful to define judgements for checking tycon signature validity separately. The judgements $\vdash \psi$ for tycon signatures, $\vdash \chi$ for opcon signatures, and $\vdash \Phi~\mathtt{sigsok}$ for all signatures in a tycon context are in Figure \ref{fig:tycon-and-opcon-signature-well-formedness}. Kind and equality kind formation are covered in the next section.

\begin{lemma}[Signatures Define Equality Type Index Kinds]
\label{lemma:sigs-define-eq-type-index-kinds}
If $\vdash \tcsig{\ktyidx}{\chi}$ then $\keq{}{\ktyidx}$.
\end{lemma}
\begin{proof} By rule induction on the assumption. The conclusion is the first premise.
\end{proof}

\begin{lemma}[Valid Opcon Structures Define Valid Opcon Signatures]
\label{lemma:chi-ok}
If $\vdash_\Phi \omega \sim \tcsig{\ktyidx}{\chi}$ then $\vdash \chi$.
\end{lemma}
\begin{proof} By rule induction on the assumption.

\pfcase{(ocstruct-intro)} We have that $\chi=\introsig{\klitidx}$ and (1) $\emptyset \vdash \ktyidx$. Apply (ocsig-intro-ok) to (1).

\pfcase{(ocstruct-targ-ok)} We have that $\chi=\chi', \opsig{\opname{op}}{\klitidx}$ and (1) $\vdash_\Phi \omega \sim \tcsig{\klitidx}{\chi'}$ and (2) $\emptyset \vdash \klitidx$. Applying the IH to (1), we have (3) $\vdash \chi'$. Apply (ocsig-targ-ok) to (3) and (2). 
\end{proof}

\begin{lemma}[Well-Defined Contexts Have Well-Formed Signatures]
\label{lemma:well-defined-contexts-have-well-formed-signatures}
If $\vdash \Phi$ then $\vdash \Phi~\mathtt{sigsok}$.
\end{lemma}
\begin{proof} By rule induction on the assumption.

\pfcase{(tcc-emp)} Apply (tcc-sigsok-emp).

\pfcase{(tcc-ext)} We have that $\Phi=\Phi', \tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\theta}$ and (1) $\vdash \Phi'$ and (2) $\keq{}{\ktyidx}$ and (3) $\vdash_\Phi \omega \sim \tcsig{\ktyidx}{\chi}$.

By the IH on (1), we have (4) $\vdash \Phi'~\mathtt{sigsok}$.

By Lemma \ref{lemma:chi-ok} on (3) we have (5) $\vdash \chi$.

By (tcsig-ok) on (2) and (5) we have (6) $\vdash \tcsig{\ktyidx}{\chi}$.

By (tcc-sigsok-ext) on (4) and (6) we have our conclusion.
\end{proof}

\begin{lemma}[Tycons Have Valid Signatures]
\label{lemma:tycons-have-valid-signatures}
If (a) $\tcdef{\tc}{\psi}{\theta} \in \Phi$ and (b) $\vdash \Phi~\mathtt{sigsok}$ then $\vdash \psi$.
\end{lemma}
\begin{proof} By rule induction on (b) and (a).

\pfcase{(tcc-sigsok-emp)} Does not apply by (a).

\pfcase{(tcc-sigsok-ext)} We have two possible cases by (a):
\begin{quote}
\pfcase{$\Phi=\Phi', \tcdef{\tc}{\psi}{\theta}$} Then the conclusion is premise 2 of (b).

\pfcase{$\Phi=\Phi', \tcdef{\tc'}{\psi'}{\theta'}$ for $\tc\neq\tc'$ and (1) $\tcdef{\tc}{\psi}{\theta} \in \Phi'$} We have (2) $\tccsigok{\Phi'}$ by premise 1 of (b). Apply the IH to (1) and (2).
\end{quote}
\end{proof}

\begin{lemma}[Tycon Context Signature Validity Respects Extension]
If (a) $\tccsigok{\Phi}$ and (b) $\tccsigok{\Phi'}$ and (c) $\text{dom}(\Phi) \cap \text{dom}(\Phi') = \emptyset$ then $\tccsigok{\Phi\Phi'}$.
\end{lemma}
\begin{proof} Rule induction on (b).

\pfcase{(tcc-sigsok-emp)} Here, $\Phi'=\emptyset$ so $\Phi\Phi'=\Phi$ and the conclusion follows by (a).

\pfcase{(tcc-sigsok-ext)} Here, $\Phi'=\Phi'', \tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}}$ and (1) $\tccsigok{\Phi''}$ and (2) $\vdash \psi$.

$\Phi\Phi''$ is well-formed because prefixing respects disjointness and (c).

Applying the IH on (a) and (1), we have that (3) $\tccsigok{\Phi\Phi''}$.

By (tcc-sigsok-ext) on (3) and (2) we have our conclusion.
\end{proof}
\newpage

% \itm & ::= & \multicolumn{1}{l}{x ~|~ \ilam{\ity}{x}{\itm} ~|~ \iap{\itm}{\itm} ~|~ \ifix{\ity}{x}{\itm} ~|~ \iLam{\alpha}{\itm} ~|~ \iAp{\ity}{\itm}}\\
% & ~|~ & \multicolumn{1}{l}{\ifold{t}{\ity}{\itm} ~|~ \iunfold{\itm} ~|~ \itriv ~|~ \ipair{\itm}{\itm} ~|~ \ifst{\itm} ~|~ \isnd{\itm}}\\
% & ~|~ & \iinl{\ity}{\itm} ~|~ \iinr{\ity}{\itm} ~|~ \icase{\itm}{x}{\itm}{x}{\itm}\\

\section{Static Language}
\subsection{Syntax}

In the paper, we described the static language (SL) as consisting of a total functional core plus three additional kinds. That core is derived directly from \cite{pfpl}. Since the form of the judgements are not technically identical, we will re-state the rules below, but we will only sketch the corresponding cases in the proofs for concision (the core ignores the tycon context, argument bound $n$, etc.).

The only change from the paper here is that we define the SL as having a built in list kind, rather than the more general mechanism for inductive kinds hinted at in the paper (lists are the only inductive kind we strictly need). The details of inductive kinds are in \cite{pfpl}, but again would distract from our main point and aren't needed to prove our theorems. Another approach is to simply allow recursive kinds, which would make it possible to define static terms that diverge. This does not break our fundamental guarantees, only decidability of typechecking (which we do not go into in detail here, as it is a more practical concern and requires non-trivial proofs of weak normalization that would also distract from our main point).

As with the IL, each choice of SL forms a dialect of @$\lambda$. We anticipate that the SL (essentially, a domain-specific language for implementing typed translation semantics), and the IL would need to evolve much more slowly than the EL (e.g. the GHC compiler's IL has remained  stable for many years), so dialect formation should be needed far more rarely % mechanized proof of our properties for the SL should begin with an existing language mechanization.

\subsection{Kinds}\label{sec:kinds}
\subsubsection{Kind Formation}

The kind formation rules are straightforward, shown in Figure \ref{fig:kind-formation}. Kind formation contexts $\kDelta$ are defined as finite mappings and thus are identified up to contraction and exchange, and we can assume the domains are disjoint. We assume substitution is defined in the usual way, so that standard substitution lemmas hold.

\begin{assumption}[Kind Variable Substitution - Kinds]
\label{lemma:kind-variable-substitution-kinds}
If $\kDelta, \kalpha \vdash \kappa'$ and $\kDelta \vdash \kappa$ then $\kDelta \vdash [\kappa/\kalpha]\kappa'$.
\end{assumption}

\begin{lemma}[Weakening of Kind Formation]
\label{lemma:weakening-of-kind-formation}
If $\kDelta \vdash \kappa$ then $\kDelta\kDelta' \vdash \kappa$.
\end{lemma}
\begin{proof} Rule induction on assumption.

\pfcases{(kf-unit), (kf-ty), (kf-ity), (kf-itm)} These rules are defined for all $\kDelta$, so we re-apply the same rule.

\pfcases{(kf-arrow), (kf-prod), (kf-sum)} Apply the IH to premise 1 to get (1) $\kDelta\kDelta' \vdash \kappa$. Apply the IH to premise 2 to get (2) $\kDelta\kDelta' \vdash \kappa_2$. Re-apply the same rule to (1) and (2).

\pfcase{(kf-list)} Here, $\kappa=\klist{\kappa'}$. Apply the IH to premise 1 to get (1) $\kDelta\kDelta' \vdash \kappa'$. Apply (kf-list) to (1).

\pfcase{(kf-alpha)} By extension of finite mappings, if $\kalpha \in \kDelta$ then (1) $\kalpha \in \kDelta\kDelta'$. Apply (kf-alpha) to (1).

\pfcase{(kf-forall)} Here, $\kappa=\kforall{\kalpha}{\kappa'}$. Apply the IH to premise 1 to get (1) $(\kDelta, \kalpha)\kDelta' \vdash \kappa'$. By exchange, $(\kDelta, \kalpha)\kDelta'=\kDelta\kDelta', \kalpha$ and thus we can apply (kf-forall) to (1).

\end{proof}

\subsubsection{Equality Kinds}

Equality kinds are included to avoid complicating type equality (and guarantee \emph{argument independence} of values, see below). Arrow kinds, polymorphic kinds and term quotation kinds (because they are argument-dependent, see below) are not equality kinds (though an approach for quantifying over only equality kinds could be introduced to allow polymorphic equality kinds, as Standard ML provides \cite{Tofte:89:TheDefinitionOfStandardML}). The rules are given in Figure \ref{fig:equality-kinds}. %We do not define an equational theory for the language at all, so this is merely meant to head off the concerns of those who might wish to develop one in the future.

\begin{lemma}[Equality Kind Well-Formedness]
\label{lemma:equality-kind-well-formedness}
If $\keq{}{\kappa}$ then $\emptyset \vdash \kappa$. 
\end{lemma}
\begin{proof} Rule induction on the assumption.

\pfcase{(keq-list)} We have $\kappa=\klist{\kappa'}$ and (1) $\keq{}{\kappa'}$. Apply IH to (1) to get that (2) $\emptyset \vdash \kappa'$. Apply (kf-list) to (2).

\pfcase{(keq-unit)} We have $\kappa=\kunit$. Apply (kf-unit).
